% STYLE NOTES
% -----------
% - Refer to the reader's colleagues as researchers or investigators?
% - Use second person?
% - Be clear that we are always ranting.

\documentclass[letterpaper]{book}
\usepackage[utf8]{inputenc}

% 1. The latex book format has pages too short
\addtolength{\topmargin}{-0.70in}
\setlength{\textheight}{2\textwidth}
% 2. If you want to scroll well, you need to center the text:
\setlength{\oddsidemargin}{3.25in}
\addtolength{\oddsidemargin}{-0.5\textwidth}
\setlength{\evensidemargin}{\oddsidemargin}
% 3. Latex is broken
\linespread{1.08}
\sloppy\sloppypar\raggedbottom\frenchspacing
% 4. Always with the headings
\renewcommand{\MakeUppercase}[1]{{#1}}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\renewcommand{\chaptermark}[1]{\markboth{\thechapter. #1}{\thechapter. #1}} 
\newcommand{\chaptermarkstar}[1]{\markboth{#1}{#1}} 

\title{\textbf{The Practice of Astrophysics}}
\author{anonymous}
\date{April 2021}

\begin{document}

\maketitle

\cleardoublepage
\tableofcontents

\chapter*{Preface}%
\chaptermarkstar{Preface}\addcontentsline{toc}{chapter}{Preface}

For the last few decades there has been a conversation going on in the astrophysics community about the disconnect between what we teach in the standard undergraduate and graduate core, and what we use in our day-to-day work as researchers.
Visualization! That's never taught to you, and many figures in many papers are hard to read.
Teaching! We are hired to teach but don't (without extra effort) obtain any training in education.
Statistics! We learn the folk activities of our advisors and supervisors, while doing research, and often wrongly.
Mentoring! We have enormous responsibilities to our students and yet we learn only by doing; that's not right, especially when things are going wrong.

These conversations about the shortcomings of our academic educations lead to enthusiastic calls to revise the curriculum and teach the things that an astrophysicist \emph{really needs to know}.
And yet: What has changed? In my institution and many others, the core requirements for the PhD (prior to admission and within the program) have barely changed in these same few decades.

People (especially parents of physics-oriented students) often ask me what is the most important thing to know or learn if you want to become a scientist.
I can surprise them by saying ``reading and writing''.
Science is the scientific literature.
When we say that ``we know the age of the Universe to better than one percent'' we mean that there is a published (and probably refereed) scientific paper that explains and justifies this point.
We produce writing.
Astrophysicists are writers.
We are also designers, coders, builders, data-processors, statisticians, visualizers, mentors, teachers, advisors, speakers, proposers, and referees.
But even these myriad activities are often or even primarily in the service of writing and publishing papers, which are the long-lasting products of our research.

In this sense, astrophysics is one of the \emph{humanities}.
It is a written literature, written by humans, recording what humans have learned, and responding to what humans want to know.

There is another category of long-lasting product, though, in addition to our writing, and that is \emph{people}:
We educate, train, mentor, and launch people into careers in academia, in schools, in national labs, in technology companies, in policy positions, and in government.
This is probably the most important part of what we do, since it combines all the challenges of humanity with ethical and moral complexities, perhaps even more extreme than in other areas because \emph{astronomy has no point in itself}.
We only do astronomy because people want to do astronomy, and want to have astronomy done.
If people lost interest in this, we would have no role in society.
There is no urgent human need for astronomy, beyond the urgent human need to understand everything that can be understood.
For these reasons---and maybe to stay consistent with Kant's categorical imperative\footnote{Kant's categorical imperative~\cite{} has multiple different versions or statements that are hard to reconcile. The statement of the imperative that I consider to be one of the few undeniably true discoveries in ethics is that people should be the \emph{ends} of our work, not the \emph{means} of our work. This principle will come up again in various places in this book, especially Chapters~??. All that said, I am not a philosopher!}---we have to make \emph{people} the focus of what we do.
We do astrophysics for people, and we do astrophysics with people, and the astrophysics we do is in the service of the lives, interests, and careers of those people.

What makes me qualified to write this book?
I have had a long career doing research in astrophysics, and teaching physics and mentoring early-career astrophysicists.
I have only done some parts of this job well.
In some sense, this book is inspired and informed by many of the mistakes I have made---mistakes in how I wrote papers, mistakes in what I put in grant proposals, mistakes in how I managed my research budget, mistakes in the classroom, and mistakes in how I mentored graduate students and postdocs.
I will try to be specific about some of these mistakes in what follows.
Some of them caused me anguish.
There is a sense in which being an academic is the greatest privilege in the world: We can work on what we want, how we want, with whoever we want.
But there is an equally true sense in which being an academic is a stressful job: We have no-one to blame but ourselves when things go wrong.
Academic astrophysics is a paradigm of the existential dystopia\footnote{In college I was taught that anguish, abandonment, and despair are the three principal existential experiences~\cite{}.}.

What will you get out of this book?
This book won't solve the fundamental problem that opens this Preface: 
After reading this book you will not know how to write, nor how to teach, nor how to mentor, nor how to perform and visualize data analyses.
This book is intended to start \emph{conversations} about these things.
It could be used (selectively I presume) in a semester-long or year-long course for graduate students and research-interested undergraduates, to supplement the standard core of classes about mechanics, thermodynamics, radiative processes, and so on.

One of the things I love about the subjects here is the following:
You will never be \emph{done} learning these things.
Some areas of human knowledge can be mastered.
For example, after 30 years, I think I know everything I will ever need to know about linear algebra.\footnote{I don't mean to be tough on linear algebra here. It is incredibly rich and rewarding subject, and it really \emph{has} taken me 30 years to learn it. I just mean that linear algebra has a clearly defined \emph{scope}, unlike, say, advising graduate students, or writing scientific abstracts.}
But I am certain that even 20 years from now---if I am still here---I will still be learning things about writing, about teaching, about advising, about making figures, about packaging software, about designing projects, and about giving talks.
That is, these subjects are unbounded in their scope and no document or project or interaction or result is ever \emph{perfect}.
(As we will see in Chapter~\ref{ch:papers}, I don't even think perfection is a goal!)
You will always have more to learn about all these things.
That makes it hard to set the scope for a book like this!
And perhaps---circling back to where we started---that makes it hard to bring these things into the curriculum of astrophysics: If we did, where would we start and where would we end?
It sure isn't obvious. And no two astronomers will agree. Here's my take.

\chapter*{Acknowledgements}%
\chaptermarkstar{Acknowledgements}\addcontentsline{toc}{chapter}{Acknowledgements}

The idea for this book came up during the pandemic years, in conversations with Adrian Price-Whelan (Flatiron). He also helped a lot, particularly on the subjects of...

If I successfully write this book, I owe Juna Kollmeier (CITA, Carnegie) a huge thanks.

Astronomical Data Group

My PhD Students. My postdocs.

My undergraduate students.

\part{Reading and writing}%

\chapter{How to read a scientific paper}\label{ch:reading}

HOGG: Note: People who teach writing often say that the best way to learn to write is to spend a lot of time reading.

\chapter{The title of your paper}\label{ch:title}

Sometimes there is a declarative sentence, like

Brainstorming words and themes.

Re-evaluating as you finish and revise the paper.

\chapter{The abstract of your paper}\label{ch:abstract}

I can't emphasize this enough: \emph{The title and abstract are the most important parts of any scientific paper}.
It is on the basis of the title and abstract that your paper gets read in the first place.
And often when an investigator is performing a literature search, they look only at the abstract to decide on the relevance of the paper.
Sometimes only the title!
For this reason, the title and abstract of a paper must be written with extreme care, and not hastily at the end of the process.

I want to emphasize this again: It is a big mistake to leave the writing of the abstract to the end of your writing process.
I see this mistake made a lot:
``How can I write my abstract? I haven't finished writing the paper. I don't know what my final conclusions are!''
This reasoning makes two mistakes:
The first is that it implies that the main point of the abstract is to communicate the \emph{results} of your work.
That's not right.
The abstract summarizes the whole paper, including context, project design, data taking, and methods, almost all of which are set quite early in the project, usually way before any writing (but see below).
The second mistake is that you don't know your results until you are done.
Okay, maybe you don't know whether you are going to get 65.1 or 65.2 for the numerical value of the Hubble constant.
But you do know that you will get \emph{some number} for that constant, and you know that it will be on the low side.

Why have I separated the writing of titles and abstracts from the writing of scientific papers, which is the subject of Chapter~\ref{ch:papers}?
Two reasons.
The first is that it is so important, it deserves its own rant.
The second is that titles and abstracts are a part of many scientific documents in addition to papers.
Grant proposals, white papers, and seminar announcements all require titles and abstracts.
If you are starting out, it is probably the case that before you ever have to write any scientific paper, you will have to write one or more titles and associated abstracts.
This is because you will give talks about your work or prepare a poster for display at a scientific meeting.

HOGG: Use the abstract writing as a way of setting project scope.
If it doesn't fit in an abstract well, the scope is too large, and the opposite.

HOGG: Final thoughts: Write your abstract first, at the project-design, conceptual stage. It helps!

\chapter{Writing scientific papers}\label{ch:papers}

HOGG: Be sure to mention that you are done when you are 99-percent done!

HOGG: Be sure to mention that you should share it \emph{on submission}, otherwise those who could help you will be too late to help you (and might even withhold their help). When you share, share openly, not selectively; you don't know who might help you.

\chapter{Dissertations and theses}

\chapter{Grant proposals, job applications}

% look at an altac_naumes tweet from 2021-04-13 about grant reviews: "the first sentence(s) of a proposal should clearly state 1) what you plan to do; 2) how you plan to do it; and 3) why the funder should care. The introduction is not a lit review.

Grant proposals are a different kind of writing entirely, since they are more explicitly \emph{persuasive} writing, and in particular, they are persuasive not about the properties of stars or galaxies, but instead about you and your research program.

HOGG: The proposal needs to be understood completely on a skim! This is hard to arrange, but here are some points.

HOGG: The proposal ought to explicitly highlight and answer the questions or talking points or language in the call for proposals.

\chapter{White papers and reports}

HOGG: Most important thing: Recommendations should be bold (or typeset obviously) and should be written so that they can be cut-and-pasted into \emph{another} report without modification.

\chapter{Typesetting and typography}

HOGG: dashes!

HOGG: Line lengths!

HOGG: Two-column is bad!

HOGG: Accessibility considerations.

\part{Teaching}

\chapter{Syllabus}

\chapter{Assignments and evaluation}

\chapter{Classroom practices}

\part{Scientific collaboration}

\chapter{Project design and design thinking}

\chapter{Authorship rules}

Thoughts:
\begin{itemize}
    \item Authorship is not an honor; it represents work.
    \item Some authors endorse the results. Do all have to?
    \item Authorship should involve significant intellectual work at least; a very simple idea is not enough to earn it.
    \item Authorship should rarely be given for work done in the deep past; at some point that work becomes dead (or it should have been published already).
    \item Author ordering depends on discipline, but we prefer models that are based on clear rules (or as clear as possible).
    \item We like models in which there is an explicit statement in the paper about who did what.
    \item PhD Theses have different rules and have to be treated differently.
\end{itemize}

\chapter{Collaboration policies}

\chapter{Open science}

\part{Meetings}

\chapter{Giving a seminar}

\chapter{How to ask a question in a seminar}

\chapter{How to answer a question in a seminar}

\chapter{Attending a scientific meeting}

\chapter{Running a scientific meeting}

\chapter{Collaboration meetings}

\part{Data analysis}

\chapter{Model building}

\chapter{Probability}

\chapter{Information theory}

\chapter{Linear algebra}

\chapter{Optimization}

\chapter{Visualization}

Figures should be readable, unambiguously on a black-and-white printer or display.
Figures should have aspect ratios such that they can be cut easily into presentation slides.
Figures should have large text such that they are readable at a glance.
Ink on figures should be used in proportion to the importance of the point being made (more ink on more important data, for example).
Data should always be dark and black-ish, models can be light and colored.
Figures should be readable by people with (at least) the most common form of color-blindness.
Lines and points should be distinguished not just by hue but also by value.
The same quantities on different plots should be plotted in the same point or line style and the same color, with the same label and same range.
And so on!

\chapter{Data structures and databases}

\chapter{Data publication and release}

\part{Software and code}

\chapter{Good practice}

\chapter{Diagnosis and debugging}

\chapter{Open-source and community}

\part{Managing and supervising}

\chapter{Supervising student projects}

\chapter{Accepting supervision; managing up}

\chapter{Mentoring}

\chapter{Equity and Inclusion}

\chapter{Hiring}

\chapter{Making and maintaining budgets}

\chapter{Project management}

\part{Reviewing the work of others}

\chapter{Referee reports}
\setcounter{page}{351}

\chapter{Letters of recommendation}

Academic appointments, from PhD positions through to faculty jobs through to promotions to tenure and full-professor status, all involve letters written by third parties.
As you proceed in astrophysics, you will have to write more and more such letters each year.
A lot of the best practices for letters are very specific to the rank of the position, the place in question, and the candidate.
But also a lot is in common to all letters.

In what follows, we will assume that you are writing for a ``candidate'' (a person you know) for a position or promotion at an ``organization.''
We will further assume that you like the candidate and you want them to succeed.
This isn't always true!
But we advise you to \emph{only write for candidates that you want to succeed}.
It is only barely ethical to agree to write a letter for someone just so you can preside over their failure.
Furthermore, your refusal to write usually conveys whatever negatives you want to convey much better than a negative letter, which is an unpleasant thing to write, an unpleasant thing to read, and generally harmful to everyone, including the writer.

So, what are the general principles for writing a good letter of recommendation?
\begin{description}
    \item[Start with a specific summary:]
    ``I am pleased to be writing on behalf of Dr. C. Andidate, who is applying for the postdoctoral fellowship at your institution. My brief summary is that Dr. Andidate is the best person in the world at finding anomalies in stellar data, having made two extremely important discoveries in this area in the last year. Andidate is already working at the level of the very best postdocs I have ever worked with, they represent a no-brainer hire for this position.''
    Your summary should say something specific about their research and say \emph{whether or not you think the organization should hire (or promote) them}.
    \item[Establish your relationship to the candidate:]
    Very briefly: How long and in what capacity have you worked with (or taught or advised) this candidate?
    Here you can also admit any possible conflict of interest, as in ``I should admit at this point that I am also a very close friend of the candidate, and I want them to succeed.''
    Sometimes you are writing for someone you have never worked with.
    In this case, say so, explicitly; also say why you are qualified to speak to the case nonetheless.
    It can work well in these cases to point out that your distance gives you ``an impartial view'' of their work and accomplishments.
    \item[Describe a bit of research, with specifics:]
    \item[Describe impact:]
    \item[Speak to the candidate's communication:]
    \item[Speak to the candidate's collaboration, mentoring, or teaching:]
    \item[Speak to the candidate's ability to raise money:]
    \item[Be specific:]
    \item[Use success-oriented language:]
    Don't say ``if you admit this student'' or ``if you invite them for an interview'' say instead ``when you admit this student'' or ``when you invite them for an interview.''
    Speak to the reader as if the reader obviously agrees with you that this person is the best.
    \item[To compare or not to compare?]
    Some letter writers feel that one should never compare one candidate to others; other writers feel that one always must.
    Our position is that any comparisons---if any---should always be specific, never vague.
    That is, ``This person is a better writer than $X$'' not ``This person is better than $X$.''
    For early-career researchers, avoid comparing at fixed status (that is, avoid comparing to others at the same career stage).
    So say ``This person is even better at theoretical dynamics than $X$ was at this same career stage,'' where $X$ is someone famous and senior to the candidate.
    Those kinds comparisons are flattering to everyone, even the senior person (because they are being held up as a standard).
    What about negative comparisons?
    Our position is that they should be softened always, as in
    ``They are less of a hard-core coder than $X$ was at this stage, but better at communicating their results,'' or
    ``They are more of a community-builder than $X$ but less of an empire builder.''
    Merit is not a one-dimensional property. There is no way that one person is \emph{better} than another, in general.
    Breathe in this important truth and write letters that are consistent with it, no matter what the assignment.
    \item[End with a specific summary:]
\end{description}
One comment to make about this structure (which involves starting and ending with a summary):
When people are in a hurry, and have to read a large number of letters, they tend to put more of their attention onto the first paragraph and the last paragraph.
This effect continues outside those terminal paragraphs:
If you want to emphasize something about the candidates research, say it \emph{first}, immediately after the opening summary.
If you want to emphasize something about the candidate's ability to raise money or get telescope time, say it \emph{last}, just before the final summary.
So make sure that any reader who reads \emph{only the beginning and end of the letter} (heavens!) would nonetheless fully understand what you think of the candidate.

Be careful about gendered and racialized language.... HOGG... Do you use titles and first names and last names equally across all your people?
If not, write about everyone as you would write about the hegemon.
Also, and obviously, use the candidate's preferred name and pronouns.
If the candidate uses ``they/them'' be careful that your sentences are absolutely unambiguous, given that.
It is better to repeat the candidate's name too much (see the first quotation in this Chapter) than it is to be clever with the grammar.

The main thing we get asked about letters is: ``How long should a letter be?''
The answer is... HOGG

All that makes sense and is easy to achieve for a candidate for whom you can say positive things about every aspect of their work.
There are some candidates that have real lacunae or deficits.
How to address these?
Our advice is to ignore them:
The letter will be positive but shorter for their lack.
If you really feel that you \emph{must} address some deficit (for example, because you were specifically asked to address it),
then do so with understanding and humility.

Sometimes you might have a collaborator who you know has been through something hard, or has some constraints, such that explaining them would violate their privacy.
It is sometimes important to mention something like this, but it is absolutely required that you do not violate their privacy.
The confidentiality of the letter of recommendation is not a license to say whatever you want.
You can't say ``this candidate is married to a trans woman, so they really want a job in a trans-friendly city.''
You can say ``for various reasons, this candidate wants to live in Washington State, so you have a good shot at recruiting them.''
You can't say ``this person got cancer during their PhD, so their productivity is all the more remarkable.''
You can say ``this person overcame very significant personal challenges during their PhD, which makes their productivity all the more remarkable.''

By the way, and out of scope: References for industrial or commercial or non-profit jobs are completely different from references for academic posiitons and promotions.
Usually they are phone calls (not letters) or questionnaires, and they don't want long answers.
They just want to know: Will this candidate steal from our organization?
The answer (we hope) is no.

Finally: A great letter violates many conventions.
If there is someone you really think is extremely good, get creative.
Be funny.
Be conversational.
Express your enthusiasm.
In part, you are trying to generate an emotional response in the reader.

\chapter{Proposal reviews}

\chapter{Tenure evaluations}

\part{Practice and principle}

\chapter{Transparency}

\chapter{Inclusion and Equity}

\chapter{Institutions and hierarchies}

\end{document}
